\pkg{cv}: An \proglang{R} Package for Cross-Validating Regression Models

### 2024-07-03 scratch/notes_john/JSS-article/JSS-article-3.Rmd
fix discussion of bias and variance in JSS paper draft

Because the $n$ models are each fit to $n - 1$ cases, LOO CV produces a nearly unbiased estimate of prediction error. The $n$ regression models are highly statistical dependent, however, based as they are on nearly the same data, and so the resulting estimate of prediction error has relatively large variance. In contrast, estimated prediction error for $k$-fold CV with $k = 5$ or $10$ (commonly employed choices) are somewhat biased but have smaller variance. It is also possible to correct $k$-fold CV for bias (see Section \ref{computation-of-the-bias-corrected-cv-criterion-and-confidence-intervals}).

>>>>

Because the $n$ models are each fit to $n - 1$ cases, LOO CV produces a nearly unbiased estimate of prediction error. The $n$ regression models are highly statistically dependent, however, based as they are on nearly the same data, and so the resulting estimate of prediction error has larger variance than if the predictions from the models fit to the $n$ data sets were independent.

Because predictions are based on smaller data sets, each of size approximately $n - n/k$, estimated prediction error for $k$-fold CV with $k = 5$ or $10$ (commonly employed choices) is more biased than estimated prediction error for LOO CV. It is possible, however, to correct $k$-fold CV for bias (see Section \ref{computation-of-the-bias-corrected-cv-criterion-and-confidence-intervals}).

The relative *variance* of prediction error for LOO CV and $k$-fold CV (with $k < n$) is more complicated: Because the overlap between the data sets with each fold omitted is smaller for $k$-fold CV than for LOO CV, the dependencies among the predictions are smaller for the former than for the latter, tending to produce smaller variance in prediction error for $k$-fold CV. In contrast, there are factors that tend to inflate the variance of prediction error in $k$-fold CV, including the reduced size of the data sets with each fold omitted and the randomness induced by the selection of folds---in LOO CV the folds are not random.

### 2024-07-03 scratch/notes_john/JSS-article/JSS-article-3.tex
fix discussion of bias and variance in JSS paper draft

Because the \(n\) models are each fit to \(n - 1\) cases, LOO CV
produces a nearly unbiased estimate of prediction error. The \(n\)
regression models are highly statistical dependent, however, based as
they are on nearly the same data, and so the resulting estimate of
prediction error has relatively large variance. In contrast, estimated
prediction error for \(k\)-fold CV with \(k = 5\) or \(10\) (commonly
employed choices) are somewhat biased but have smaller variance. It is
also possible to correct \(k\)-fold CV for bias (see Section
\ref{computation-of-the-bias-corrected-cv-criterion-and-confidence-intervals}).

@@

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 969978
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.18252
bias-adjusted cross-validation criterion = 19.17193
95% CI for bias-adjusted CV criterion = (15.72359, 22.62027)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> BayesRule
\end{CodeInput}
\begin{CodeOutput}
function(y, yhat) {
  if (!all(y %in% c(0, 1)))
    stop("response values not all 0 or 1")
  if (any(yhat < 0) ||
      any(yhat > 1))
    stop("fitted values outside of interval [0, 1]")
  yhat <- round(yhat)
  result <- mean(y != yhat) # proportion in error
  attr(result, "casewise loss") <- "y != round(yhat)"
  result
}
<bytecode: 0x13b56c108>
<environment: namespace:cv>
\end{CodeOutput}
\end{CodeChunk}

>>>>

Because the \(n\) models are each fit to \(n - 1\) cases, LOO CV
produces a nearly unbiased estimate of prediction error. The \(n\)
regression models are highly statistically dependent, however, based as
they are on nearly the same data, and so the resulting estimate of
prediction error has larger variance than if the predictions from the
models fit to the \(n\) data sets were independent.

Because predictions are based on smaller data sets, each of size
approximately \(n - n/k\), estimated prediction error for \(k\)-fold CV
with \(k = 5\) or \(10\) (commonly employed choices) is more biased than
estimated prediction error for LOO CV. It is possible, however, to
correct \(k\)-fold CV for bias (see Section
\ref{computation-of-the-bias-corrected-cv-criterion-and-confidence-intervals}).

The relative \emph{variance} of prediction error for LOO CV and
\(k\)-fold CV (with \(k < n\)) is more complicated: Because the overlap
between the data sets with each fold omitted is smaller for \(k\)-fold
CV than for LOO CV, the dependencies among the predictions are smaller
for the former than for the latter, tending to produce smaller variance
in prediction error for \(k\)-fold CV. In contrast, there are factors
that tend to inflate the variance of prediction error in \(k\)-fold CV,
including the reduced size of the data sets with each fold omitted and
the randomness induced by the selection of folds---in LOO CV the folds
are not random.

@@

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 431467
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.192
bias-adjusted cross-validation criterion = 19.18107
95% CI for bias-adjusted CV criterion = (15.71858, 22.64356)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> BayesRule
\end{CodeInput}
\begin{CodeOutput}
function(y, yhat) {
  if (!all(y %in% c(0, 1)))
    stop("response values not all 0 or 1")
  if (any(yhat < 0) ||
      any(yhat > 1))
    stop("fitted values outside of interval [0, 1]")
  yhat <- round(yhat)
  result <- mean(y != yhat) # proportion in error
  attr(result, "casewise loss") <- "y != round(yhat)"
  result
}
<bytecode: 0x13fb26ea8>
<environment: namespace:cv>
\end{CodeOutput}
\end{CodeChunk}
